learning how to mutate 

```{r}
primary_24 %>% 
  mutate(new_column = votes - election_day) 
#using columns instead of rows! votes and election_day are both columns, and when we do this it adds a new column with this data for allll the candidates
#give your new column a good name 
#you can use this to do percentages! 
mutate(percent = (percent equation)) makes a new column 
#you can still group by and summarize after mutating!
#you can use mutate to make things all caps: str_to_upper(column), str_to_lower(column), str_to_title(column)
```

time for case_when 
used to classify 

```{r}
mutate(
  amount_category = case_when (
    amount < 100 ~ "less than $100", 
    amount >= & amount < 500 ~ "between $100 and $499", 
    amount >= 3000 ~ "$3000 or more "
  )
)
#"amount" here is a column name 
```

filter()

```{r}
maryland_winred_contributions %>% 
  filter(amount_category == "less than $100")
#now we have a dataset with just the less than 100 rows 
#and we can add stuff
maryland_winred_contributions %>% 
  filter(amount_category == "less than $100")
group_by(last_name, first_name) %>% 
  summarise(total_given = sum(amount))
blah blah etc etc 

```
if it's in a column, you can filter for it 
remember to use == in filter 
you can filter for multiple things: & is and and | is or 

lubridate to change dates from <chr> to <date>

```{r}
library(lubridate)

maryland_expenses %>% 
  select(expenditure_date) %>% 
  mutate (new_expenditure_date = mdy(expenditure_date))
#match to what the og data was -- mdy is if the og data comes in m/d/yy format. dmy and ymd are also things. 
```


When you do summarize(), make sure you're making a new column. 
group_by(column_name) %>% 
summarize(whatever_the_fuck = n())


"yeah welcome to my hell"

select(column_a, column_b) to display just those columns 

```{r}
%in%
```
```{r}
md_voters_2024 %>% 
 mutate(
   region = case_when(
     COUNTY %in% c("one", "two", "three",) ~ "Capital Region"
   )
# %in% c() makes a list!


TRUE ~ "other region"
# this is saying: for things other than what we listed, call it "region" 
```

cleaning data using case_when
```{r}
md_voters_2024 %>% 
  mutate(
    county = case_when(
      county == "Saint Mary's" ~ "St. Mary's", 
      TRUE ~ COUNTY 
    )
  )
#used to change the spelling of St Marys 
```

janitor: clean_names
```{r}
md_voters_2024 %>% 
  clean_names()

#you absolutely should use it all the time
#import data, then clean names 
```

check for problems 
problems(data_set_name)

rename(newname = oldname)

mutate(amount = as.numeric(amount))
#changes to a number

get_dupes() 
#finds duplicates

mutate(zip = str_sub(zip, start=1L, end=5L))

mutate(city = str_to_title(city))





