---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
options(scipen=999)
library(refinr)
library(janitor)
library(dplyr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
donors <- read_csv("data/donors.csv")

donors <- donors %>% 
  mutate(
    amount = str_remove_all(amount, "\\$"),
    amount = str_remove_all(amount,",")
  )
glimpse(donors)

donors <- donors %>% 
  mutate(date = mdy(date)) %>% 
  mutate(amount = as.numeric(amount)) %>% 
  mutate(zip = str_sub(zip, start = 1L, end = 5L)) %>% 
  mutate(committee_zip = str_sub(committee_zip, start = 1L, end = 5L))

glimpse(donors)

donors <- donors %>% 
  rename(
    name_clean = "name-clean"
  )
glimpse(donors)

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")


# Clean required data and prepare for analysis if needed. 



```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Add in the week before session began, and the two businessmen gave more than $87,000 to Republican legislators and the North Carolina Republican Party, a key cog in the campaign machine this year as the two major parties fight for control of the General Assembly in November's elections.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

#week before session is jan 3-9
donors %>% 
  filter(name_clean == "GEORGE LEWIS STROBEL II" | name_clean == "ROBIN E DELMER") %>% 
  group_by(name_clean) %>% 
  summarise(amount = sum(amount)) %>% 
  arrange(desc(amount))

61250+57250
  
donors %>% 
  filter(name_clean == "GEORGE LEWIS STROBEL II" | name_clean == "ROBIN E DELMER") %>% 
  group_by(committee_name) %>% 
  summarise(amount = sum(amount)) %>% 
  arrange(desc(amount))
#rabon = dem
#michael lee = repub = 12500	
#saine = repub = 12500	
#hise = repub = 12500	
#jackson = repub = 10000	
#moore = repub = 10000	
#brown = repub = 10000	
#berger = repub = 10000	
#nc rec = real estate commission = 6000
#randleman = repub = 5000	
#szoka = repub = 4000	
#brawley = repub = 2500	
#lewis = repub = 2500	 (later got in legal trouble for corruption lmao)
#tillman = repub = 2500	
#bell = repub = 2500	
#burr = repub = 1000

12500+12500+12500+10000+10000+10000+10000+5000+4000+2500+2500+2500+2500+1000 = 97500

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-