---
title: "Reverse Engineering Project"
author: Rachel, Fiona, Brinkley, Yugo
date: 3-12-25"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
options(scipen=999)
library(refinr)
library(janitor)
library(dplyr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
donors <- read_csv("data/donors.csv")

donors <- donors %>% 
  mutate(
    amount = str_remove_all(amount, "\\$"),
    amount = str_remove_all(amount,",")
  )
glimpse(donors)

donors <- donors %>% 
  mutate(date = mdy(date)) %>% 
  mutate(amount = as.numeric(amount)) %>% 
  mutate(zip = str_sub(zip, start = 1L, end = 5L)) %>% 
  mutate(committee_zip = str_sub(committee_zip, start = 1L, end = 5L))

glimpse(donors)

donors <- donors %>% 
  rename(
    name_clean = "name-clean"
  )
glimpse(donors)

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")


# Clean required data and prepare for analysis if needed. 



```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Add in the week before session began, and the two businessmen gave more than $87,000 to Republican legislators and the North Carolina Republican Party, a key cog in the campaign machine this year as the two major parties fight for control of the General Assembly in November's elections.”
* **Analysis summary**: We were able to confirm these findings. Once we filtered, grouped_by, and summarized, we were able to figure out what kinds of donations they were making and to who. 

```{r}
# Put code to reverse engineer sentence here

#week before session is jan 3-9
donors %>% 
  filter(name_clean == "GEORGE LEWIS STROBEL II" | name_clean == "ROBIN E DELMER") %>% 
  group_by(name_clean) %>% 
  summarise(amount = sum(amount)) %>% 
  arrange(desc(amount))

61250+57250

  donors %>% 
  filter(name_clean == "GEORGE LEWIS STROBEL II" | name_clean == "ROBIN E DELMER") %>% 
  group_by(committee_name) %>% 
  summarise(amount = sum(amount)) %>% 
  arrange(desc(amount))
#rabon = dem
#michael lee = repub = 12500	
#saine = repub = 12500	
#hise = repub = 12500	
#jackson = repub = 10000	
#moore = repub = 10000	
#brown = repub = 10000	
#berger = repub = 10000	
#nc rec = real estate commission = 6000
#randleman = repub = 5000	
#szoka = repub = 4000	
#brawley = repub = 2500	
#lewis = repub = 2500	 (later got in legal trouble for corruption lmao)
#tillman = repub = 2500	
#bell = repub = 2500	
#burr = repub = 1000

12500+12500+12500+10000+10000+10000+10000+5000+4000+2500+2500+2500+2500+1000 = 97500
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: “Senate President Pro Tem Phil Berger's campaign alone raised nearly $222,000 during session. Twenty-three people gave him the maximum allowed: $5,200.”

* **Analysis summary**: We were able to confirm the finding! When we filtered for Berger's campaign, the dates of the short session, and the amount given (5200), we got 23 rows. This number proves the sentence above that states 23 individual donors (organizations or people) donated the max amount allowed. 

```{r}
# Put code to reverse engineer sentence here

berger <- donors %>% 
  filter(date >= '2018-05-16' & date <= '2018-06-29') %>% 
  filter(committee_name == "PHILIP E BERGER COMM") %>% 
  filter(amount == 5200) %>% 
  group_by(name_clean, date, amount) %>% 
  summarise(sum(amount))
#https://www.sierraclub.org/north-carolina/blog/2018/07/legislature-2018-wrapup is where we got the ending date!



# Display results of code below this codeblock


```

### Sentence 3

* **Sentence text**: Beer distributors also gave during session. Mark Craig, the president of distributor R.H. Barringer, put more than $30,000 into state legislative campaigns during the six-week period, including the maximum $5,200 to Berger. He was the top individual donor to legislative campaigns during session.
* **Analysis summary**: Yes, we were able to confirm these findings. There was a little confusion on the 6-week session for the second part,  but everything else went fine. 

```{r}
# Put code to reverse engineer sentence here
Mark_Craig <- donors|>
  filter(date >= '2018-05-16' & date <= '2018-06-29') |>
  filter(name_clean == "MARK R CRAIG")|>
  filter(amount == 5200) |>
  group_by(name_clean, date, amount) |>
  summarize(sum(amount))

#above proves max 5200 donation to Berger (run it one line at a time)

donors %>% 
  filter(name_clean == "MARK R CRAIG") %>% 
  select(name_clean, date, amount) %>% 
  arrange(date)

#above proves most of his donations were during legislative session (not all, though!) 

donors %>% 
  filter(name_clean == "MARK R CRAIG") %>% 
  group_by(name_clean) %>% 
  summarize(
    total_amount = sum(amount, na.rm = TRUE)
  )
#na.rm comes from chatgpt
  

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: “Add in donations to the two main political parties and to a handful of other large fundraising vehicles for state legislative races, and the in-session giving topped $3.4 million. Gov. Roy Cooper's main campaign account brought in another $407,000 during session.”

* **Analysis summary**: We were able to find the total in-session giving of over $3.4 million. Our number was 3446455. As of right now, we're unable to prove the Cooper thing -- we can't find his campaign in the data. 

```{r}
# Put code to reverse engineer sentence here
#step 1: find total in-session giving. 

donors %>% 
  filter(date >= '2018-05-16' & date <= '2018-06-29') %>% 
  summarize(sum(amount))

#step 2: find out how much money Gov. Roy Cooper's campaign account brought in. 
campaign_totals <- donors %>% 
  filter(date >= '2018-05-16' & date <= '2018-06-29') %>% 
  group_by(committee_name) %>% 
  summarize(sum = sum(amount)) %>% 
  filter(sum >= 407000)

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: “A WRAL News analysis of campaign finance filings also turned up $16,000 in donations to legislative campaigns during session from 15 PACs with a lobbyist. Each would appear to be a violation of state fundraising rules, but spokespeople for most of the PACs, or the companies they represent, said there was a simple explanation: The checks were written before session.”
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
#find PACs 
# Filter for donations from PACs during the legislative session
pac_donations <- donors |>
  filter(date >= '2018-05-16' & date <= '2018-06-29') |>
  filter(str_detect(name_clean, "PAC"))  # More flexible matching

# View total amount
total_pac_amount <- pac_donations |>
  summarise(total_amount = sum(amount))

print(total_pac_amount)

# Optional: View individual PAC donations (to check for lobbyist links manually or join with a lobbyist list)
print(pac_donations)

# Display results of code below this codeblock

#USED CHATGPT -- HELPED IMPROVE ORIGINAL CODE. BRINKLEY HAD THE CODE WRITTEN, GOT AN ERROR, ASKED CHAT, CHAT ADDED THE PART FOR TOTAL PAC DONATIONS. 

```

-30-