---
title: "lab_02"
author: "derek willis"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters. Which counties have seen the most? Take a look at county population figures here and compare the two in your answer: https://worldpopulationreview.com/us-counties/maryland.

**A1**
There were over 17000 new voters on or after 2024-07-21. The county with the most new voters was Montgomery, which makes sense because according to the site it's the most populous county in MD. Other states with high population and high numbers of new voters included Baltimore city/county, PG county, and anne arundel. 

```{r}
new_voters <- md_voters_2024 %>% mutate(since_721 = case_when(
  COUNTY_REGISTRATION_DATE >= "2024-07-21" ~ "new"
))
```

```{r}
only_new_voters <- new_voters %>% filter(since_721 == "new")
```

```{r}
only_new_voters %>% 
  group_by(COUNTY) %>% 
  summarize(county = n()) %>% 
  arrange(desc(county))
```


**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2**
July 21-Aug 7 were different from the rest because weeks 30 and 31 (July 21-Aug 3) had a really high number of registrations (breaking 7000 both weeks). Week 24 earlier in the summer had the highest number of registrations, though. I looked it up and all I could find was Hunter Biden being convicted -- not sure why that week had such a high number of registrations. 

```{r}
library(lubridate)

actual_week_number_data <- md_voters_2024 %>% 
  mutate(week_number = week(COUNTY_REGISTRATION_DATE))

actual_week_number_data %>% 
  group_by(week_number) %>% 
  summarize(n()) %>% 
  arrange(desc(week_number))
#OK so it looks like the weeks with the highest voter registration were 31, 30, 28, and 24. 30 and 31 are the last 2 weeks of July starting with 7-21, so that makes sense. The other 2 were in June and early July. 
```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?
**A3** 
It looks like July had the largest number of registrations, which makes sense because it had that 2-week stretch of like 14000 new voters after Biden dropped out. 

```{r}
md_voters_2024 %>% 
  mutate(month = months(COUNTY_REGISTRATION_DATE)) %>% 
  group_by(month) %>% 
  summarize(month = n()) %>% 
  arrange(desc(month))
#looking it up. 
#OHHH RIGHT FLOOR DATE 
```
```{r}
md_voters_2024 %>% 
  mutate(month = floor_date(COUNTY_REGISTRATION_DATE)) %>% 
  group_by(month) %>% 
  summarize(month = n())
#HOW DO I DO THIS. checking notes. 
#dammit i didn't write this down. 
#consulting the internet. 
#rdocumentation.org save me
#trying again. 
```
```{r}
md_voters_2024 %>% 
  mutate(month = floor_date(
    COUNTY_REGISTRATION_DATE, 
    unit = "month")) %>% 
  group_by(month) %>% 
  summarize(month = n())
#holy cow i think that worked
#thank you R help window, you're more useful than the internet. 
```



**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are. Which regions and parties account for the most new voters?

**A4** 
Now, they won't give me the name of the region when I group by and summarize. So if I assume it's in alphabetical order, central MD had the most new voters. Also wouldn't label the data when I tried to figure out party registration. I am going to take the L on this. 
I talked to chatgpt for help with case_when and commented our conversation below. 

```{r}
region_voters <- md_voters_2024 %>% 
  mutate(
    region = case_when(
      COUNTY ==  "Allegany" ~ "Western", 
      COUNTY ==  "Garrett" ~ "Western", 
      COUNTY ==  "Washington" ~ "Western", 
      COUNTY == "Frederick" ~ "Capital",
      COUNTY == "Montgomery" ~ "Capital",
      COUNTY == "Prince George's" ~ "Capital", 
      COUNTY == "Anne Arundel" ~ "Central",
      COUNTY == "Baltimore City" ~ "Central", 
      COUNTY == "Baltimore County" ~ "Central",
      COUNTY == "Carroll" ~ "Central", 
      COUNTY == "Harford" ~ "Central", 
      COUNTY == "Howard" ~ "Central", 
      COUNTY == "Calvert" ~ "Southern", 
      COUNTY == "Charles" ~ "Southern", 
      COUNTY == "Saint Mary's" ~ "Southern", 
      COUNTY == "Kent" ~ "Eastern Shore", 
      COUNTY == "Queen Anne's" ~ "Eastern Shore", 
      COUNTY == "Talbot" ~ "Eastern Shore", 
      COUNTY == "Caroline" ~ "Eastern Shore", 
      COUNTY == "Dorchester" ~ "Eastern Shore", 
      COUNTY == "Wicomico" ~ "Eastern Shore", 
      COUNTY == "Somerset" ~ "Eastern Shore", 
      COUNTY == "Worcester" ~ "Eastern Shore", 
      COUNTY == "Cecil" ~ "Eastern Shore")
  )

#I GOT STUCK IMMEDIATELY AND CONSULTED CHAT GPT. CHAT GPT HELPED. COMMENTING IN OUR CONVERSATION. 


#Hey dude, I'm working in R right now on some data that includes a column listing the county of each instance. I'm supposed to use case_when to add a new "region" column, classifying each instance by the region each county is in. But when I do code like region = case_when(       COUNTY =  Allegany ~ "Western", R gives me an error and says it fails to classify the first side of the formula. Any idea how I can use case_when on character strings?
#ChatGPT said:
#Hey! I see the issue here. In case_when, when you're comparing strings (like county names), you need to use == for comparison, not =. The = operator is for assignment in R, whereas == is for equality comparison.

#Try this corrected syntax:

#r
#Copy
#library(dplyr)

#df <- df %>%
 # mutate(region = case_when(
  #  COUNTY == "Allegany" ~ "Western",
   # COUNTY == "Montgomery" ~ "Central",
    #COUNTY == "Baltimore" ~ "Central",
    #TRUE ~ "Other"  # Default case for counties not explicitly listed
  #))
#The key change is using == instead of = for string comparison. Also, TRUE ~ "Other" is a catch-all for counties that don't match any of the conditions you've listed.

#Let me know if that works!


#You said:
#ARE YOU SERIOUS

#ChatGPT said:
#Haha, yeah, I’m serious! It’s one of those little quirks in R that can trip you up. Using == for comparison instead of = is a common mistake, but once you get the hang of it, it’ll be smooth sailing.

#Let me know how it goes! If you run into any other issues, I’ve got your back!


#You said:
#Ok I changed it to == but still got an error message. Here's the code, do you know why it can't find the character string?
#md_voters_2024 %>% 
 # mutate(
    #region = case_when(
     # COUNTY ==  Allegany ~ "Western"
    #)
 # )
#ChatGPT said:
#Gotcha! I see what's going on here. The error likely happens because the string "Allegany" is not being recognized due to the way it's referenced. In R, character strings need to be in quotes (i.e., "Allegany") to be interpreted properly. It seems like Allegany is being treated as a variable name rather than a string.

#Here’s the corrected version of your code:

#r
#Copy
#md_voters_2024 %>% 
  #mutate(
    #region = case_when(
      #COUNTY == "Allegany" ~ "Western"
    #)
 # )
#Notice the "Allegany" is now inside quotes. That should fix the issue!

#Let me know if it works now!


#You said:
#DUUUUDDDEEE THAT WORKS! 

#And then we had a conversation about AI not being as scary as I thought.
```

```{r}
region_voters %>% 
  group_by(region, PARTY) %>% 
  summarize(region = n(), PARTY = n())
```


**Q5** What's an additional question you could ask of this data?

**A5** 
It would be interesting to compare historical voter registration (maybe from the 2020 and 2016 elections) for each region and party to voter registration for 2024. Is there more voter registration now than there used to be in certain regions and for certain parties? And what percentage of registered voters in each region voted in the last 3 elections? Has voter turnout changed over the past decade or so? 
